\textls[-0]{\thetitle}

### 2023-03-18 cc-condensed.tex
question
adding back the bit about how linkages
Fix canonicity model, reply comments.
edits

\noindentparagraph{Consistency.}
One of the most fundamental properties of a dependent type theory is consistency.
\begin{theorem}[Consistency]
\label{thm:consistency}
  The typing judgment $\goodTerm{\cdot}{t}{\bot}$ is not derivable for any term~$t$.
\end{theorem}

@@

\noindentparagraph{Canonicity.}

@@

We prove \cref{thm:consistency,thm:canonicity} by constructing a logical-relations model
for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,kaposi2019gluing,sterling2019algebraic}.
%
The model interprets the bottom type~$\bot$ as an empty set, from which
\cref{thm:consistency} follows.
A closed, well-formed type $\goodType{\cdot}{T}{}$ is interpreted as
a logical predicate on closed terms: the predicate includes all
closed, ``reducible'' terms of type~$T$. So \cref{thm:canonicity} follows.
%\EDJ{I will add a quote around reducible here. it is my bad to introduce "reducible" without defining it first. But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}\EDJreply{Because in the operational semantic, using ``reducible'' is totally fine. But in declarative style, only ``canonicity'' this word is making sense. But this ``canonicity'' is not friendly to the audience not familar with declartive style.}
% The proof is available as a supplemental file in Agda syntax.
The construction of the logical-relations model is available in
\ifreport
\cref{sec:canonicity-model}.
\else
an appendix.
\fi

>>>>

\noindentparagraph{Consistency and canonicity.}
One of the most fundamental properties of a dependent type theory is consistency.
\begin{theorem}[Consistency]
\label{thm:consistency}
  The typing judgment $\goodTerm{\cdot}{t}{\bot}$ is not derivable for any term~$t$.
\end{theorem}

@@

@@

We prove \cref{thm:consistency,thm:canonicity} by constructing a logical-relations model
for the well-formedness rules, following prior approaches~\cite{coquand2018canonicity,kaposi2019gluing,sterling2019algebraic}.
%
The model interprets a linkage~$\LkgAdd{\lkg}{t}$ as a pair where the
second component is a function (modeling late binding),
as rule \ruleref{l/add} indicates.
%
The model interprets the bottom type~$\bot$ as an empty set, from which
\cref{thm:consistency} follows.
A closed, well-formed type $\goodType{\cdot}{T}{}$ is interpreted as
a logical predicate on closed terms: the predicate includes all
closed, ``reducible'' terms of type~$T$.
\cref{thm:canonicity} follows from this interpretation.
%\EDJ{I will add a quote around reducible here. it is my bad to introduce "reducible" without defining it first. But my intension is to give a feeling of the dynamic operational semantic for the reader (of course OS is not here in our setting)}\YZreply{Is there a more accurate way to say things here?}\EDJreply{Because in the operational semantic, using ``reducible'' is totally fine. But in declarative style, only ``canonicity'' this word is making sense. But this ``canonicity'' is not friendly to the audience not familar with declartive style.}
% The proof is available as a supplemental file in Agda syntax.
The construction of the logical-relations model is available in
\ifreport
\cref{sec:canonicity-model}.
\else
an appendix.
\fi\YZ{and as fake Agda sources?}

### 2023-03-18 metatheory-condensed.tex
edits

For accessibility of the main text, \cref{sec:metatheory2} presents \TT using
named variables and meta-level substitution.
For clarity of proof details, the technical appendix supplementing
\cref{sec:metatheory2} instead uses de Bruijn indices and explicit substitution.
%
We acknowledge that the presentation is dense for an audience
unacquainted with MLTT, so we summarize the salient points first.

@@

The definitive version containing all the rules in \TT is stated in
\ifreport \cref{sec:complete-fmltt}.
\else a supplemental appendix.
\fi
using de Bruijn indices and explicit substitution.
An effort to mechanize the formalization is underway, by following  \citet{altkap2016} 
to use quotient inductive-inductive types as 
the logical framework to state the syntax and using Agda's
\lsti{REWRITE} pragma to replace the limited uses of equality reflection appeared in 
an extensional meta\-logic.

>>>>

For accessibility of the main text, \cref{sec:metatheory2} presents \TT using
named variables and meta-level substitution.
For clarity of proof details, a technical appendix supplementing
\cref{sec:metatheory2} instead uses \citeauthor{debruijn1964} indices and explicit substitution \cite{abadi1989subst}.
%
We acknowledge that the presentation is dense for an audience
unacquainted with MLTT, so we summarize the salient points first.

@@

The definitive version containing all the rules in \TT is stated in
\ifreport \cref{sec:complete-fmltt}.
\else a technical appendix.
\fi
This definitive version uses de Bruijn indices and explicit substitution.
An effort to mechanize the formalization is underway. It follows \citet{altkap2016}
in using quotient inductive-inductive types (QIITs) as 
the logical framework to state intrinsically typed syntax (and also
in using Agda's \lsti{REWRITE} pragma to replace the limited uses of equality
reflection, as Agda does not support QIITs natively).
%in the extensional meta\-logic.
